---
  title: "1.0_tests"
author: "Andrea Estandia<br>"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  github_document:
  toc: true
toc_depth: 4
editor_options:
  chunk_output_type: console
---
  <br>
  
```{r setup, echo=FALSE}
# Knitr settings: 
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
options(scipen = 999)
```

```{r, echo=FALSE}
source("./src/ae-candidate-gene-analysis_source.R")
```

Read allelic richness data and Ho
```{r}
ar.c <-read_csv("data/allelic-richness-candidate.csv")
ho.c <-read_csv("data/allelic-richness-candidate.csv")
```

T-test for differences between clusters
```{r}
list_genes <- colnames(ar.c[,c(3:6)])

purrr::map(ar.c[3:6], ~t.test(.x ~ ar.c$archipelago))
purrr::map(ho.c[3:6], ~t.test(.x ~ ho.c$archipelago))
```

Read data and create subset by gene

```{r, echo=FALSE}
df <-
  read_csv(file.path(data_path, "20211001_df_cand_AE+SMC+BR.csv")) %>%
  mutate_if(is.character, as.factor)

CREB1 <- df %>%
  drop_na(short_creb1)

ADCYAP <- df %>%
  drop_na(short_adcyap)

CLOCK <- df %>%
  drop_na(short_clock)

NPAS2 <- df %>%
  drop_na(short_npas2)
```

ANOVA and Tukey tests for DRD4 and SERT
```{r}
list_variants <- colnames(df[,grep("DRD|SERT", colnames(df))])
subset_variants <- df[,grep("DRD|SERT", colnames(df))]

list_variants_out <- list()
for (variant in list_variants){
  tmp <- df %>%
    dplyr::select(name, population, variant, locator) %>%
    drop_na(variant)
  print(summary(aov(as.formula(paste0(as.character(variant), "~ population")), data = tmp)))
  print(TukeyHSD(aov(as.formula(paste0(as.character(variant), "~ population")), data = tmp)))
}
```

Summary stats, ANOVA and Tukey tests for candidate microsatellite
```{r}
list_variants <- colnames(df[,grep("mean_adcyap|mean_npas|mean_clock|mean_creb", colnames(df))])
subset_variants <- df[,grep("mean_adcyap|mean_npas|mean_clock|mean_creb", colnames(df))]

list_variants_out <- list()
for (variant in list_variants){
  tmp <- df %>%
    dplyr::select(name, population, variant, locator) %>%
    drop_na(variant)
  print(summary(aov(as.formula(paste0(as.character(variant), "~ population")), data = tmp)))
  print(TukeyHSD(aov(as.formula(paste0(as.character(variant), "~ population")), data = tmp)))
}

for (variant in list_variants){
  print(group_by(df, population) %>%
          summarise(
            count = n(),
            mean = mean(get(variant), na.rm = TRUE),
            sd = sd(get(variant), na.rm = TRUE)
          ))
}
```

Create subset with only Australia, Tasmania and Tasmanian migrants removing those birds that weren't caught in winter. Test whether there are differences between migrants, Australians and Tasmanian individuals
```{r}
migratory.subset_total <- df %>%
  filter(population %in% c("Australia","Tasmania")) %>%
  dplyr::arrange(population) %>%
  mutate(date2 = dmy(date)) %>% 
  mutate(month=month(date2)) %>% 
  filter(((population=="Australia" & month >= 4 & month <= 10 ))|
          migrant == "yes" |
          (population == "Tasmania" & month == 6)) %>% 
  mutate(population==as.character(population)) %>% 
  mutate(population=ifelse(migrant=="yes", "Tasmania migrants", as.character(population))) %>% 
  mutate(population=ifelse(population=="Tasmania", "Tasmania non-migrants", population)) %>% 
  mutate(population==as.factor(population))
   

clock.migr.aov <- aov(mean_clock ~ population, data = migratory.subset_total)
clock.migr.tukey <- TukeyHSD(clock.migr.aov)

list_variants <- colnames(migratory.subset_total[,grep("DRD|SERT|mean_adcyap|mean_clock|mean_creb1|mean_npas2", colnames(migratory.subset_total))])
subset_variants <- migratory.subset_total[,grep("DRD|SERT|mean_adcyap|mean_clock|mean_creb1|mean_npas2", colnames(migratory.subset_total))]

list_variants_out <- list()
for (variant in list_variants){
  tmp <- migratory.subset_total %>%
  dplyr::select(name, population, variant, locator) %>%
  drop_na(variant)
  #print(summary(aov(as.formula(paste0(as.character(variant), "~ population")), data = tmp)))
  print(variant)
  print(TukeyHSD(aov(as.formula(paste0(as.character(variant), "~ population")), data = tmp)))
}
```

Pairwise t-test for allele length

```{r, echo=FALSE}
pairwise.t.test(CREB1$mean_creb1, CREB1$population, p.adj = 'bonf')
pairwise.t.test(CLOCK$mean_clock, CLOCK$population, p.adj = 'bonf')
pairwise.t.test(ADCYAP$mean_adcyap, ADCYAP$population, p.adj = 'bonf')
pairwise.t.test(NPAS2$mean_npas2, NPAS2$population, p.adj = 'bonf')
```

T-test to know whether partial migrants and the rest of the populations have a significant different mean CLOCK value
```{r, echo=FALSE}
#some stats to explore the data
group_by(CREB1, migrant) %>%
  summarise(
    count = n(),
    mean = mean(mean_creb1, na.rm = TRUE),
    sd = sd(mean_creb1, na.rm = TRUE)
  )

#Create subsets to do the t-test
migrant <- df[which(df$migrant == 'yes'), ]
nonmigrant <- df[which(df$migrant == 'no'), ]

#t-test
t.test(migrant$mean_creb1, nonmigrant$mean_creb1)
```