---
  title: "1.0_model_migrants"
author: "Andrea Estandia<br>"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  github_document:
  toc: true
toc_depth: 4
editor_options:
  chunk_output_type: console
---
  <br>
  
```{r setup, echo=FALSE}
# Knitr settings: 
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
options(scipen = 999)
```

```{r, echo=FALSE}
source("./src/ae-candidate-gene-analysis_source.R")
```

Read data and create keep only CLOCK data

```{r, echo=FALSE}
CLOCK <-
  read_csv(file.path(data_path, "20211001_df_cand_AE+SMC+BR.csv")) %>%
  mutate_if(is.character, as.factor) %>% 
  drop_na(mean_clock)
```


Create subset with only Australia, Tasmania and Tasmanian migrants removing those birds that weren't caught in winter. Test whether there are differences between migrants, Australians and Tasmanian individuals

```{r}
migratory.subset_total <- CLOCK %>%
  filter(population %in% "Tasmania") %>%
  mutate(date2 = dmy(date)) %>% 
  mutate(month=month(date2)) %>% 
  filter(((population=="Tasmania" & state=="Queensland" & month >= 4 & month <= 10 ))|
          migrant == "yes" |
          (population == "Tasmania" & state=="Tasmania" & month >=4 & month <=10)) %>% 
  mutate(population==as.character(population)) %>% 
  mutate(population=ifelse(migrant=="yes", "Tasmanian migrants", as.character(population))) %>% 
  mutate(population=ifelse(population=="Tasmania", "Tasmanian non-migrants", population)) %>% 
  mutate(population==as.factor(population)) %>% 
  filter(migrant!="NA")
  
```

Run a frequentist approach to explore difference in means across populations
```{r}
list_variants <- colnames(migratory.subset_total[,grep("DRD|SERT|mean_adcyap|mean_clock|mean_creb1|mean_npas2", colnames(migratory.subset_total))])
subset_variants <- migratory.subset_total[,grep("DRD|SERT|mean_adcyap|mean_clock|mean_creb1|mean_npas2", colnames(migratory.subset_total))]

list_variants_out <- list()
for (variant in list_variants){
  tmp <- migratory.subset_total %>%
  dplyr::select(name, population, variant, locator) %>%
  drop_na(variant)
  #print(summary(aov(as.formula(paste0(as.character(variant), "~ population")), data = tmp)))
  print(variant)
  print(TukeyHSD(aov(as.formula(paste0(as.character(variant), "~ population")), data = tmp)))
}
```


A Bayesian approach: Do partial migrants and the rest of the populations have a significant different mean CLOCK value?
```{r}
migratory.subset <- 
  migratory.subset_total %>% 
  filter(population != "Australia")

get_prior(mean_clock ~ population,
  data = migratory.subset)

bprior <- prior(normal(0,20), class = b)

mod <- brm(
  mean_clock ~ population,
  data = migratory.subset,
  prior = bprior)

migratory_raw <- migratory.subset %>% 
  ggplot(
    aes(
    x = fct_reorder(population, mean_clock),
    y = mean_clock,
    color = population,
    fill = population
  )
) +
  geom_dotplot(
    binaxis = 'y',
    stackdir = 'center',
    binwidth = 0.12,
    dotsize = 1.5
  ) +
  labs(x = "", y = "Mean CLOCK length (bp)\n") + 
  #stat_summary(fun.data = data_summary, color ="black") + 
  scale_x_discrete(breaks=c("Australia", "Tasmanian non-migrants", "Tasmanian migrants"),
                   labels =c("Australia", "Tasmanian non-migrants", "Tasmanian migrants"))+
  scale_fill_manual(values = c("grey", "grey")) + 
  scale_color_manual(values = c("grey", "grey")) +
  theme_minimal() + 
  theme(legend.position = "none",
        panel.grid.major = element_blank()) 


migratory_posterior <- migratory.subset %>%
  data_grid(population) %>%
  add_epred_draws(mod) %>%
  ggplot(aes(x = .epred, y = population)) +
  stat_dotsinterval(quantiles = 100)+
  theme_minimal() +
  theme(
    legend.position = "none",
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.text.x = element_text(hjust=1, size = text_size),
    axis.title.y = element_blank(),
    axis.title = element_text(size = text_size),
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(color = "#424242"))+
  labs(
    x = "\nCLOCK mean length (bp)",
    title = "Posterior estimates of the group means"
  )

migratory_raw + migratory_posterior+plot_annotation(tag_levels = 'A')
plot(hypothesis(mod, "Intercept = Intercept + populationTasmaniannonMmigrants"))
```
