---
  title: "1.0_model_migrants"
author: "Andrea Estandia<br>"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  github_document:
  toc: true
toc_depth: 4
editor_options:
  chunk_output_type: console
---
  <br>
  
```{r setup, echo=FALSE}
# Knitr settings: 
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
options(scipen = 999)
```

```{r, echo=FALSE}
source("./src/ae-candidate-gene-analysis_source.R")
```

Read data and create keep only CLOCK data

```{r, echo=FALSE}
CLOCK <-
  read_csv(file.path(data_path, "20211001_df_cand_AE+SMC+BR.csv")) %>%
  mutate_if(is.character, as.factor) %>% 
  drop_na(mean_clock)
```


Create subset with only Australia, Tasmania and Tasmanian migrants removing those birds that weren't caught in winter. Test whether there are differences between migrants, Australians and Tasmanian individuals

```{r}
migratory.subset_total <- df %>%
  filter(population %in% c("Australia","Tasmania")) %>%
  dplyr::arrange(population) %>%
  mutate(date2 = dmy(date)) %>% 
  mutate(month=month(date2)) %>% 
  filter(((population=="Australia" & month >= 4 & month <= 10 ))|
          migrant == "yes" |
          (population == "Tasmania" & month == 6)) %>% 
  mutate(population==as.character(population)) %>% 
  mutate(population=ifelse(migrant=="yes", "Tasmania migrants", as.character(population))) %>% 
  mutate(population=ifelse(population=="Tasmania", "Tasmania non-migrants", population)) %>% 
  mutate(population==as.factor(population))
  
```

Run a frequentist approach to explore difference in means across populations
```{r}
list_variants <- colnames(migratory.subset_total[,grep("DRD|SERT|mean_adcyap|mean_clock|mean_creb1|mean_npas2", colnames(migratory.subset_total))])
subset_variants <- migratory.subset_total[,grep("DRD|SERT|mean_adcyap|mean_clock|mean_creb1|mean_npas2", colnames(migratory.subset_total))]

list_variants_out <- list()
for (variant in list_variants){
  tmp <- migratory.subset_total %>%
  dplyr::select(name, population, variant, locator) %>%
  drop_na(variant)
  #print(summary(aov(as.formula(paste0(as.character(variant), "~ population")), data = tmp)))
  print(variant)
  print(TukeyHSD(aov(as.formula(paste0(as.character(variant), "~ population")), data = tmp)))
}
```


A Bayesian approach: Do partial migrants and the rest of the populations have a significant different mean CLOCK value?
```{r}
migratory.subset <- 
  migratory.subset_total %>% 
  filter(population != "Australia")

mod <- brm(
  mean_clock ~ population,
  data = migratory.subset
)

migratory.subset %>%
  data_grid(population) %>%
  add_fitted_draws(mod) %>%
  ggplot(aes(x = .value, y = population, fill=population)) +
  stat_dotsinterval(quantiles = 100)+
  scale_fill_manual(values = c("#ccb416", "#538699"))+
  theme_minimal() +
  theme(
    legend.position = "none",
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    axis.text.x = element_text(hjust=1, size = text_size),
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    axis.title = element_text(size = text_size),
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(color = "#424242"))+
  labs(
    x = "CLOCK mean lenght(bp)",
    title = "Posterior estimates of the group means"
  )
```
